Parameter estimation

### 2023-04-20 SFE-article/Sections/Gouverning_equation_derivation.tex
08.03
09.03
10.03
16.03
20.04

	Let's assume that any properties of the flow are uniform across any given cross-section of an extractor. The variation of the cross-section might be an result of partial filling of an extractor or its irregular shape. In reality, such a flow is two-dimensional, because with the area changing as a function of $z$, in actuality there will be flow-field variations in both directions. The assumption of quasi-one-dimensional flow dictates that the flow properties are function of $z$ only. The equations described by quasi-one-dimensional assumption hold: (1) mass conservation, (2) Newton's second law, and (3) energy conservation. To ensure that these physical principles are satisfied the modified governing equation can be derived.	Let's start with the integral form of the continuity equation:

@@

	Above partial differential equation form of the continuity equation suitable for unsteady, quasi-one-dimensional flow. It ensures that mass is conserved for this mode of the flow. The $A_f(z)$ is an arbitrary function, which describe change of the cross-section of an extractor. The function $A_f(z)$ can be defined as $A_f(z) = \textbf{A} \varepsilon(z)$, where $\varepsilon$ is the bed porosity and $\textbf{A}$ is the cross-section of an empty extractor.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \textbf{A} \varepsilon (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \textbf{A} \varepsilon (z) V \right)}{\partial z} = 0
		\end{equation}
	}
	The equation can be simplified by cancel out a constant $\textbf{A}$

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \varepsilon (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \varepsilon (z) V \right)}{\partial z} = 0
		\end{equation}
	}

	If so called superficial velocity is defined $u=\varepsilon V$, the mass continuity becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \varepsilon (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f u \right)}{\partial z} = 0
			\label{EQ: Mass_Continuity}
		\end{equation}
	}

@@

	The transport of a chemical species, in this case a solute, can be described by analogous equation to the Eq. \ref{EQ: Continuity_integral_general} with additional terms on the right-hand side. The first term on the right-hand side describes that a substance goes from high density regions to low density regions and is based on the Fick's law ${\footnotesize\left( J_{diff} = D\cfrac{\partial C_f}{\partial z} \right)}$. The other term correspond for the mass transfer between solid and fluid phases, which is treated as a source term.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_f} C_f d\mathcal{V}_f + \iint_S C_f \textbf{V} \cdot \textbf{dS} = \iint_S J_{diff} \cdot \textbf{n}~\textbf{dS} + \cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_s} C_s d\mathcal{V}_s
			\label{EQ: Integral_Mass_Species_general}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{align}
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_f} C_f d\mathcal{V}_f &= \cfrac{\partial}{\partial t}\left( C_f A_f dz \right) \\
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_s} C_s d\mathcal{V}_s &= \cfrac{\partial}{\partial t}\left( C_s A_s dz \right)
		\end{align}
	}

@@

	{\footnotesize
		\begin{equation}
			\iint_S C_f \textbf{V} \cdot \textbf{dS} = C_f V dA_f + C_f A_f dV + V A_f dC_f 
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( C_f A_f \right)}{\partial t} + \cfrac{\partial \left( C_f A_f V \right)}{\partial x} = \cfrac{\partial \left( C_s A_s \right) }{\partial t} + \cfrac{\partial \left( J_{diff} A_f \right) }{\partial z}
		\end{equation}
	}

	By defining $A_f = A \cdot \varepsilon$, $A_s = A \cdot \left( 1-\varepsilon \right)$ and $u=V \cdot \varepsilon$, and assuming that $A$ is constant, the above equation becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( C_f \varepsilon \right)}{\partial t} + \cfrac{\partial \left( C_f u\right)}{\partial x} = \cfrac{\partial \left( C_s (1-\varepsilon) \right) }{\partial t} + \cfrac{\partial \left( J_{diff} \varepsilon \right) }{\partial z}
		\end{equation}
	}

	By expanding above equation, splitting variable and assuming that ${\footnotesize \cfrac{\partial \varepsilon}{\partial t}=0}$ we get

	{\footnotesize
		\begin{equation}
			\cfrac{\partial C_f}{\partial t} + \cfrac{u}{\varepsilon} \cfrac{\partial C_f}{\partial z} + \cfrac{C_f}{\varepsilon} \cfrac{\partial u}{\partial z} = \cfrac{1-\varepsilon}{\varepsilon} \cfrac{\partial C_s}{\partial t} + \cfrac{D}{\varepsilon} \cfrac{\partial C_f}{\partial z} \cfrac{\partial \varepsilon}{\partial z} + \cfrac{\partial}{\partial z} \left( D \cfrac{\partial C_f}{\partial z} \right)
		\end{equation}
	}

	The equation can be further simplified if $\cfrac{\partial u}{\partial z} = \cfrac{\partial \varepsilon}{\partial z} = D = 0$, which corresponds to the assumptions of constant velocity along the bed( which might be a case of isothermal and low-mach number flow), constant porosity( which comes from the assumption of constant area for both solid and fluid phase) and no radial diffusion.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial C_f}{\partial t} + \cfrac{u}{\varepsilon} \cfrac{\partial C_f}{\partial z}  = \cfrac{1-\varepsilon}{\varepsilon} \cfrac{\partial C_s}{\partial t} 
			\label{EQ: Simplified_Species_Conservation}
		\end{equation}
	}

@@

	The Eq. \ref{EQ: Conservative_Momentum} can be expanded further by assuming that $A_f = \textbf{A}\varepsilon$

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f V \textbf{A} \varepsilon \right) }{\partial t} + \cfrac{\partial \left( \rho_f V^2 \textbf{A} \varepsilon \right) }{\partial z} = - 	\textbf{A} \varepsilon \cfrac{\partial P}{\partial t}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f V \varepsilon \right) }{\partial t} + \cfrac{\partial \left( \rho_f V^2 \varepsilon \right) }{\partial z} = - \varepsilon 	\cfrac{\partial P}{\partial t}
		\end{equation}
	}

	If the superficial velocity $u=\varepsilon V$ is introduced, then the momentum conservation becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f u \right)}{\partial z} + \cfrac{\partial \left( \rho_f u^2/\varepsilon \right) }{\partial z} = -\varepsilon \cfrac{\partial 	P}{\partial z}
		\end{equation}
	}

@@

	The Eq. \ref{EQ: Non-conservative_Momentum} is stylistically the same as the general momentum conservation for one-dimensional flow with no-body forces. The momentum equation can be expressed in terms of superficial velocity $u=V\varepsilon$.

	{\footnotesize
		\begin{equation}
			\rho_f \cfrac{\partial \left(u / \varepsilon \right)}{\partial t} + \rho_f \cfrac{u}{\varepsilon} \cfrac{\partial \left( u / \varepsilon \right)}{\partial z} = - \cfrac{\partial P}{\partial z}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\varepsilon} \cfrac{\partial u}{\partial t} + \rho_f u \cfrac{\partial \varepsilon^{-1}}{\partial t} + \rho_f \cfrac{u}{\varepsilon} \cfrac{1}{\varepsilon} \cfrac{\partial u}{\partial z} + \rho_f \cfrac{u}{\varepsilon} u \cfrac{\partial \varepsilon^{-1} }{\partial z} = - \cfrac{\partial P}{\partial z}
		\end{equation}
	}

	If the bed is not compressible and doesn't change its properties during the batch, then ${\footnotesize\cfrac{\partial \varepsilon}{\partial t}=0}$

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\varepsilon} \left( \cfrac{\partial u}{\partial t} + \cfrac{u}{\varepsilon} \cfrac{\partial u}{\partial z} + u^2 \cfrac{\partial \varepsilon^{-1}}{\partial z} \right) = -\cfrac{\partial P}{\partial z}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\varepsilon} \left( \cfrac{\partial u}{\partial t} + \cfrac{u}{\varepsilon} \cfrac{\partial u}{\partial z} \right) = -\cfrac{\partial P}{\partial z}
			\label{EQ: Non-conservative_Momentum_Const_Porosity}
		\end{equation}
	}

@@

	{TODO: \color{red}finish derivation of energy equation for real gas and introduce homogenous version of the energy equation}

>>>>

	Following the work of \citet{Anderson1995}, the governing equations for compressible fluid with non-uniform cross-section can be obtained. Let's assume that any properties of the flow are uniform across any given cross-section of an extractor. The variation of the cross-section might be an result of partial filling of an extractor or its irregular shape. In reality, such a flow is two-dimensional, because with the area changing as a function of $z$, in actuality there will be flow-field variations in both directions. The assumption of quasi-one-dimensional flow dictates that the flow properties are function of $z$ only. The equations described by quasi-one-dimensional assumption hold: (1) mass conservation, (2) Newton's second law, and (3) energy conservation. To ensure that these physical principles are satisfied the modified governing equation can be derived.	Let's start with the integral form of the continuity equation:

@@

	Above partial differential equation form of the continuity equation suitable for unsteady, quasi-one-dimensional flow. It ensures that mass is conserved for this mode of the flow. The $A_f(z)$ is an arbitrary function, which describe change of the cross-section of an extractor. The function $A_f(z)$ can be defined as $A_f(z) = \textbf{A} \phi(z)$, where $\phi$ is the bed porosity and $\textbf{A}$ is the cross-section of an empty extractor.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \textbf{A} \phi (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \textbf{A} \phi (z) V \right)}{\partial z} = 0
		\end{equation}
	}
	The equation can be simplified by cancel out a constant $\textbf{A}$

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \phi (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f \phi (z) V \right)}{\partial z} = 0
		\end{equation}
	}

	If so called superficial velocity is defined $u=\phi V$, the mass continuity becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f \phi (z) \right)}{\partial t} + \cfrac{\partial \left( \rho_f u \right)}{\partial z} = 0
			\label{EQ: Mass_Continuity}
		\end{equation}
	}

@@

	The transport of a chemical species, in this case a solute, can be described by analogous equation to the Eq. \ref{EQ: Continuity_integral_general} with additional terms on the right-hand side. The first term on the right-hand side describes that a substance goes from high density regions to low density regions and is based on the Fick's law $\left( J_{diff} = D_e^M\frac{\partial c_f}{\partial z} \right)$. The other term correspond for the mass transfer between solid and fluid phases, which is treated as a source term.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_f} c_f d\mathcal{V}_f + \iint_S c_f \textbf{V} \cdot \textbf{dS} = \iint_S J_{diff} \cdot \textbf{n}~\textbf{dS} + \cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_s} c_s d\mathcal{V}_s
			\label{EQ: Integral_Mass_Species_general}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{align}
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_f} c_f d\mathcal{V}_f &= \cfrac{\partial}{\partial t}\left( c_f A_f dz \right) \\
			\cfrac{\partial}{\partial t} \iiint_{\mathcal{V}_s} c_s d\mathcal{V}_s &= \cfrac{\partial}{\partial t}\left( c_s A_s dz \right)
		\end{align}
	}

@@

	{\footnotesize
		\begin{equation}
			\iint_S c_f \textbf{V} \cdot \textbf{dS} = c_f V dA_f + c_f A_f dV + V A_f dc_f 
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( c_f A_f \right)}{\partial t} + \cfrac{\partial \left( c_f A_f V \right)}{\partial z} = \cfrac{\partial \left( c_s A_s \right) }{\partial t} + \cfrac{\partial \left( J_{diff} A_f \right) }{\partial z}
		\end{equation}
	}

	By defining $A_f = A \cdot \phi$, $A_s = A \cdot \left( 1-\phi \right)$ and $u=V \cdot \phi$, and assuming that $A$ is constant, the above equation becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( c_f \phi \right)}{\partial t} + \cfrac{\partial \left( c_f u\right)}{\partial z} = \cfrac{\partial \left( c_s (1-\phi) \right) }{\partial t} + \cfrac{\partial \left( J_{diff} \phi \right) }{\partial z}
		\end{equation}
	}

	By assuming that ${\footnotesize \frac{\partial \phi}{\partial t}=0}$ and expanding $J_{diff}$, we get

	{\footnotesize
		\begin{equation}
			\cfrac{\partial c_f}{\partial t} + \cfrac{1}{\phi} \cfrac{\partial \left( c_f u\right)}{\partial z} = \cfrac{ (1-\phi)}{\phi} \cfrac{\partial c_s }{\partial t} + \cfrac{1}{\phi} \cfrac{\partial}{\partial z} \left( D_e^M \cfrac{\partial c_f}{\partial z} \right)
		\end{equation}
	}

	The equation can be further simplified if $\frac{\partial u}{\partial z} = \frac{\partial \phi}{\partial z} = D_e^M = 0$, which corresponds to the assumptions of constant velocity along the bed( which might be a case of isothermal and low-mach number flow), constant porosity( which comes from the assumption of constant area for both solid and fluid phase) and no radial diffusion.

	{\footnotesize
		\begin{equation}
			\cfrac{\partial c_f}{\partial t} + \cfrac{u}{\phi} \cfrac{\partial c_f}{\partial z}  = \cfrac{1-\phi}{\phi} \cfrac{\partial c_s}{\partial t} 
			\label{EQ: Simplified_Species_Conservation}
		\end{equation}
	}

@@

	The Eq. \ref{EQ: Conservative_Momentum} can be expanded further by assuming that $A_f = \textbf{A}\phi$

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f V \textbf{A} \phi \right) }{\partial t} + \cfrac{\partial \left( \rho_f V^2 \textbf{A} \phi \right) }{\partial z} = - 	\textbf{A} \phi \cfrac{\partial P}{\partial t}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f V \phi \right) }{\partial t} + \cfrac{\partial \left( \rho_f V^2 \phi \right) }{\partial z} = - \phi 	\cfrac{\partial P}{\partial t}
		\end{equation}
	}

	If the superficial velocity $u=\phi V$ is introduced, then the momentum conservation becomes

	{\footnotesize
		\begin{equation}
			\cfrac{\partial \left( \rho_f u \right)}{\partial z} + \cfrac{\partial \left( \rho_f u^2/\phi \right) }{\partial z} = -\phi \cfrac{\partial 	P}{\partial z}
		\end{equation}
	}

@@

	The Eq. \ref{EQ: Non-conservative_Momentum} is stylistically the same as the general momentum conservation for one-dimensional flow with no-body forces. The momentum equation can be expressed in terms of superficial velocity $u=V\phi$.

	{\footnotesize
		\begin{equation}
			\rho_f \cfrac{\partial \left(u / \phi \right)}{\partial t} + \rho_f \cfrac{u}{\phi} \cfrac{\partial \left( u / \phi \right)}{\partial z} = - \cfrac{\partial P}{\partial z}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\phi} \cfrac{\partial u}{\partial t} + \rho_f u \cfrac{\partial \phi^{-1}}{\partial t} + \rho_f \cfrac{u}{\phi} \cfrac{1}{\phi} \cfrac{\partial u}{\partial z} + \rho_f \cfrac{u}{\phi} u \cfrac{\partial \phi^{-1} }{\partial z} = - \cfrac{\partial P}{\partial z}
		\end{equation}
	}

	If the bed is not compressible and doesn't change its properties during the batch, then ${\footnotesize\cfrac{\partial \phi}{\partial t}=0}$

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\phi} \left( \cfrac{\partial u}{\partial t} + \cfrac{u}{\phi} \cfrac{\partial u}{\partial z} + u^2 \cfrac{\partial \phi^{-1}}{\partial z} \right) = -\cfrac{\partial P}{\partial z}
		\end{equation}
	}

@@

	{\footnotesize
		\begin{equation}
			\cfrac{\rho_f}{\phi} \left( \cfrac{\partial u}{\partial t} + \cfrac{u}{\phi} \cfrac{\partial u}{\partial z} \right) = -\cfrac{\partial P}{\partial z}
			\label{EQ: Non-conservative_Momentum_Const_Porosity}
		\end{equation}
	}

@@

%	{TODO: \color{red}finish derivation of energy equation for real gas and introduce homogenous version of the energy equation}
